A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations

### 2017-07-07 Rotation paper/draft_working.tex
save

Visibility-based pipelines are not fundamentally limited from using partially redundant baselines. In fact, most sensitivity forecasts to date do include partial redundancy. Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in sky-tracking measurements. Arrays capable of such tracking include MWA and LOFAR. Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA-low.

@@

Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
the source, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.

@@

We begin with the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
\end{aligned}
\label{eq:Vis1}
\end{equation}
where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline vector, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. The beam power pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.

@@

\caption{Example beam response of HERA (left) and PAPER (right) antennas. Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith. Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. \label{fig:Beam}}
\end{figure}

@@

Delayed-transformed visibility has gained popularity in recent years since it has been shown that foregrounds are isolated in delay space. We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform} as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.

@@

We can rewrite the delayed-transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}.

Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets. 
The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
in the topocentric frame with the 3 dimensional rotation operator $\Gamma$:
\small
\[
\begin{aligned}V(\boldsymbol{b'},\tau') & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}+\tau'\right)\nu_{r}\right]\end{aligned}.
\]
\normalsize
With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rA^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],
\end{aligned}
\label{eq:main}
\end{equation}
where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center. \footnote{Turns out for drift-scan measurements, the first-order interpretation is largely sufficient, but we keep the term for now to reserve the option to determine its exact form numerically.} In transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
\begin{equation}
\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)+i\psi_{\nu}\right].
\end{equation}
\end{widetext}
The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field, and the first inequality follows from the assumption that
the 3D power spectrum varies negligibly over the $k$-space of interest. The careful reader may have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$. In the flat-sky limit for example, such a relation would take the form $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$. Instrumental chromaticity and foreground isolation will be examined in \ref{sec:chromaticity}. Since $\Gamma$
is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$.

@@

\begin{equation}
\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right].
 \end{aligned}
\label{eq:final}
\end{equation}

In other words the power spectrum estimate from visibilities of a baseline pair is given by 
\begin{equation}
 P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight is defined as 
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
with

\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu}
\]

Notice that $\Theta$ has no dependence on $\tau$. In Eq. \eqref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}. We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu_tracking}
\]

@@

To first order, the physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers. By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time. When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith. The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
\[
\psi_{\nu}\sim 2\pi\Delta\tau\nu.
\]
In Fig. \ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq. \eqref{eq:Thetanu} and Eq. \eqref{eq:Thetanu_tracking}). In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$ns. In the tracking case, because of minimal zenith movement, only second order effects are observed. The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do. We refer the reader to Appendix \ref{sec:appA} for further explanation. In both cases, the full effects are encapsulated in the phase of $\Theta_\nu(\psi_\nu=0)$ and can be thus determined empirically.

@@

To illustrate the effect of rephasing, we compare in Fig. \ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz. The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above. The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}. We see that although the amplitude of correlations match up for all time offsets, the phase would only locally match. The need for rephasing can thus be understood as a spatial decoherence effect. Summing over frequency without proper rephasing leads to destructive interference and signal loss. The wider the frequency profile, the more destructive the interference would be. Only after rephasing to the correct time-offset for each baseline pairs can we constructively combine the frequency channels.

@@

\caption{Comparisons of the peak phases of two different frequencies. Specifically shown are the real parts of $\Theta_{\nu}$. First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents. The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.055 sidereal days. The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks. }
\label{fig:freqdiff}
\end{figure}

Once rephased, the overall weight can be constructed by $\Theta=\int \Theta_{\nu}d\nu$, and thus
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|e^{i\psi_{\nu}}\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
\label{eq:final_new}
\end{equation}

@@

So far for clarity and generality we have avoided the traditional formulation of radio astronomy in terms of the $uv$-plane. To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions. For this section, we relax into the flat sky approximation and examine a single frequency of $150$MHz.

@@

\begin{figure*}[h!]
\includegraphics[width=\textwidth]{sensitivity_1}

>>>>

Visibility-based pipelines are not fundamentally limited from using partially redundant baselines. In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}. Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in sky-tracking measurements. Arrays capable of such tracking include MWA and LOFAR. Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA-low.

@@

Radio interferometric observations are often described in the coordinates $uvw$ define as:
\[
(u, v, w) = \frac{\nu}{c}\boldsymbol{b}, 
\]
where $\boldsymbol{b}$ is the baseline vector in Cartesian coordinates, with first and second coordinates pointing East and North, respectively.  $\nu$ is the frequency of observation, and $c$ is the speed of light. 
Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.

@@

We begin with the visibility as commonly defined in the literature (e.g.
\citealt{TMS, first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
\end{aligned}
\label{eq:Vis1}
\end{equation}
where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A(\hat{\boldsymbol{s}},\nu)$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. The beam power pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.

@@

\caption{Example beam response of HERA (left) and PAPER (right) antennas. Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. \label{fig:Beam}}
\end{figure}

@@

Delayed-transformed visibility has gained popularity in recent years since it has been shown that foregrounds are isolated in delay space. We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform} as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right),\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.

@@

We can rewrite the delayed-transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}-\tau\right)\nu_{r}\right],
\end{aligned}
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}.

Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets. 
The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
in the topocentric frame with the 3 dimensional rotation operator $\Gamma$:

\[
\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\\
& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
\]
where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center. \footnote{Turns out for drift-scan measurements, the first-order interpretation is largely sufficient, but we keep the term for now to reserve the option to determine its exact form numerically.}

With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rA^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],
\end{aligned}
\label{eq:main}
\end{equation}
In transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
\begin{equation}
\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)+i\psi_{\nu}\right].
\end{equation}
\end{widetext}
The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field, and the first inequality follows from the assumption that
the 3D power spectrum varies negligibly over the $k$-space of interest. The careful reader may have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$. In the flat-sky limit for example, such a relation would take the form $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$. Instrumental chromaticity and foreground isolation will be examined in \ref{sec:chromaticity}. Since $\Gamma$
is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$.

@@

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)\\
 & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
 \end{aligned}
\label{eq:final}
\end{equation}

In other words the power spectrum estimate from visibilities of a baseline pair is given by 
\begin{equation}
 \hat{P}(k_{b,\tau}) \equiv \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight is defined as 
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
with

\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{-i\psi_{\nu}}.
\label{eq:Thetanu}
\]

Notice that $\Theta$ has no dependence on $\tau$. In Eq. \eqref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}. We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{-i\psi_{\nu}}.
\label{eq:Thetanu_tracking}
\]

@@

To first order, the physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers. By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time. When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith. The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
\[
\psi_{\nu}\sim 2\pi\Delta\tau\nu.
\]
In Fig. \ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq. \eqref{eq:Thetanu} and Eq. \eqref{eq:Thetanu_tracking}). In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$ns. In the tracking case, because of minimal zenith movement, only second order effects are observed. The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do. We refer the reader to Appendix \ref{sec:appA} for further explanation. In both cases, the full effects are encapsulated in the phase of $\Theta_\nu$ and can be thus determined empirically without any additional computation.

@@

To illustrate the effect of rephasing, we compare in Fig. \ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145 GHz and 0.155 GHz. The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above. The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}. Summing over frequency without proper rephasing leads to destructive interference and signal loss. The wider the frequency profile, the more destructive the interference would be. Only after rephasing to the correct time-offset for each baseline pairs can we constructively combine the frequency channels.

We see from Fig. \ref{fig:freqdiff} that although the amplitude of correlations match up for all time offsets, the phase would only locally match. The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center would not extend to the entire beam pattern (Fig. \ref{fig:beamfringe}). Thus while integrating over frequencies and sky-direction in Eq. \eqref{eq:Theta} there are necessarily sky-directions that do not add coherently. This decoherence across the beam pattern cannot be removed and plays a fundamental role in determining how much sensitivity one can recover from nearly equivalent baselines.

@@

\caption{Comparisons of the peak phases of two different frequencies. Specifically shown are the real parts of $\Theta_{\nu}$. First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents. The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.0325 sidereal days. The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks. }
\label{fig:freqdiff}
\end{figure}

@@

So far for clarity and generality we have avoided the traditional formulation of radio astronomy in terms of the $uv$-plane. To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions. For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$MHz.

@@

\begin{figure*}[h!]
\includegraphics[width=0.95\textwidth]{sensitivity_1}
